---
title: "Week5"
author: "Scott Reed"
date: "9/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyr)
library(dplyr)
library(ggplot2)
```

#### Airline CSV

What writes great csv files R can read without a problem? R does. So let us create a matrix (transposed because I misread things), and write that out. We then read it back in.
```{r airline}
airlineData<-t(matrix(c("","","Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle","Alaska", "on time", 497,221,212,503,1841,"","delayed",62,12,20,102,305,"","","","","","","","AM WEST","on time",694,4840,383,320,201,"","delayed",117,415,65,129,61),nrow=7))
write.csv(airlineData,"airlineData.csv")
airlineData <- read.csv("airlineData.csv", as.is = TRUE)
kable(airlineData)
```

## Smashed Data. 

First we take off the left index count and get a data frame. Then we take the top row and make the the names of the data. 
Next we cleanup the blank row. Then we dupe the category level labels.

```{r cleanup}

airlineData<-as.data.frame(airlineData[-1],stringsAsFactors = FALSE)
names(airlineData)<-airlineData[1,]
names(airlineData)[1:2]<-c("Airline","Flight Status")
airlineData<-airlineData[c(-1,-4),]
airlineData[2,1]<-airlineData[1,1]
airlineData[4,1]<-airlineData[3,1]
kable(airlineData)
```

Because it says to use tidyr we can't just use the base functions quixotically. So we gather into City with Flights excluding airlinge and flight status and sorting on airline.
```{r tidy}
airlineDataLong <- gather(airlineData ,key=City, value="Flights", -"Flight Status",-Airline) %>% arrange(Airline)
airlineDataLong$Flights<-as.numeric(airlineDataLong$Flights)
kable(airlineDataLong)
```


### Analysis

```{r ontimepercentage}
onTime <- airlineDataLong %>% spread(`Flight Status`,"Flights")
onTime$onTimePercent <- onTime$`on time` / (onTime$delayed+ onTime$`on time`) *100
onTime <-onTime %>% arrange(onTimePercent)
ggplot(onTime, aes(City, onTimePercent, color=Airline, size=40))+geom_point()
```

We can clearly see that for ever city Am West has a lower on-time percentage. 