---
title: "Week9"
author: "Scott Reed"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(jsonlite)
library(keyring)
library(lubridate)
library(purrr)
library(RCurl)
library(stringi)
library(tidyr)
library(ggplot2)
nytApiKey <- key_list("NYTAPI")[[2]]
nytApiSecret<-key_get("NYTAPI",nytApiKey)
```

## Fetching Data

Let's fetch bestselling authors monthly for a year

```{r}
startDate<-lubridate::make_date(2018,1,1)
dates<-startDate + months(0:11)
```


We can get the hardcover fiction list, and the ebook fiction list with our api key, but must add a wait period. 
```{r }

hardCoverLists<-purrr::map(dates, function(x) {
Sys.sleep(30) 
jsonlite::fromJSON(paste( "https://api.nytimes.com/svc/books/v3/lists/",x,"/hardcover-fiction.json?api-key=",nytApiKey,sep=""))
}
)
```

```{r }

youngAdult<-purrr::map(dates, function(x) {
Sys.sleep(30) 
jsonlite::fromJSON(paste( "https://api.nytimes.com/svc/books/v3/lists/",x,"/young-adult-hardcover.json?api-key=",nytApiKey,sep=""))
}
)
```


#### Analysis


With this we can see that YA has more weeks on list, and both lists have a lot of new enterants (the mean number of weeks on list goes way down) in the last quarter. 

```{r}
weeksOnList <- data.frame(Month=month(dates,label = TRUE),ya=unlist(purrr::map(youngAdult, ~ mean(.x$results$books$weeks_on_list))),adult=unlist(purrr::map(hardCoverLists, ~ mean(.x$results$books$weeks_on_list))))
weeksOnList %>% gather("BookType","weeks", -Month )

weeksOnList %>% gather("BookType","weeks", -Month ) %>% ggplot(aes(Month,weeks,group=BookType,color=BookType))+ geom_line()
```

We can also see how many have official NYTimes reviews. It looks like for our time period this doesn't include YA. 

```{r}
bookReviewRatio <- data.frame(Month=month(dates,label = TRUE),ya=unlist(purrr::map(youngAdult, ~ mean(1-stri_isempty( .x$results$books$book_review_link)))),adult=unlist(purrr::map(hardCoverLists, ~ mean(1-stri_isempty( .x$results$books$book_review_link)))))

bookReviewRatio %>% gather("BookType","HaveReview", -Month )

bookReviewRatio %>% gather("BookType","HaveReview", -Month ) %>% ggplot(aes(Month,HaveReview,group=BookType,color=BookType))+ geom_line()

```