---
title: "Data607-Assignment2"
author: "Scott Reed"
date: "9/5/2019"
output:
  pdf_document: default
  html_document: default
---

## Database and R

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RPostgreSQL)
library(knitr)
```

***********

*Note:* Please configure to a blank postgres db (use: createdb DBNAME) with the proper connection strings 

***********

### Postgress connection

```{r connection, ECHO=FALSE, include=FALSE}

dbuser <- "postgres"
dbpass <- "REPLACEME"
dbhost <- "localhost"
dbname <- "REPLACEME"
dbport <- 5432
drv <- dbDriver("PostgreSQL")
conn <- dbConnect(drv, host=dbhost, port=dbport,user=dbuser, password=dbpass)
```

## Data table setup

We use three data tables one movies, one for raters and then a join table for ratings. 

```{r data Tables setup}
dbExecute(conn, "CREATE TABLE IF NOT EXISTS movies(
    movie_id  INT GENERATED ALWAYS AS IDENTITY PRIMARY KEY,
    movie_name VARCHAR NOT NULL
);")
dbExecute(conn, "CREATE TABLE IF NOT EXISTS raters(
    rater_id  INT GENERATED ALWAYS AS IDENTITY PRIMARY KEY,
    rater_name VARCHAR NOT NULL
);")
dbExecute(conn, "CREATE TABLE IF NOT EXISTS movieratings(
    rating_id  INT GENERATED ALWAYS AS IDENTITY PRIMARY KEY,
    rater_id INT NOT NULL REFERENCES raters(rater_id),
    movie_id INT NOT NULL REFERENCES movies(movie_id),
    rating decimal
);")

```

### Table setup

#### Movies

| field      | type       |
|------------|------------|
| movie_id   | int (auto) |
| movie_name | varchar    |


#### Raters

| field      | type       |
|------------|------------|
| rater_id   | int (auto) |
| rater_name | varchar    |

#### MovieRatings

| field      | type       |
|------------|------------|
| rating_id  | int (auto) |
| rater_id   | int (fk)   |
| movie_id   | int (fk)   |
| rating     | decimal    |

### get some names

We shall grab some name data from NYC and sample a number of them
```{r get names}
names <- read.csv("https://data.cityofnewyork.us/api/views/25th-nujf/rows.csv?accessType=DOWNLOAD")
someNames <- sample(as.character(names[,4]),50)
head(someNames)
```

We then write them to the database, and read it back for the IDs. 

```{r insert raters}
dbBegin(conn)
tblRaters <- as.data.frame(someNames)
names(tblRaters) <- c("rater_name")
dbExecute(conn,sqlAppendTable(conn, "raters", tblRaters,row.names = FALSE ))
dbCommit(conn)
tblRaters = dbReadTable(conn,"raters")
head(tblRaters)
```

### Get some movies

We get some movies sample them and load them into a database table much like names

```{r}
load(url("https://stat.duke.edu/~mc301/data/movies.Rdata"))
someMovies <- sample(as.character(movies$title),6)
head(someMovies)
dbBegin(conn)
tblMovies <- as.data.frame(someMovies)
names(tblMovies) <- c("movie_name")
dbExecute(conn,sqlAppendTable(conn, "movies", tblMovies,row.names = FALSE ))
dbCommit(conn)
tblMovies = dbReadTable(conn,"movies")
kable(tblMovies)
```

### Add some ratings

To generate some ratings we first cross apply movie_id and rater_id and then provide ratings up to 6. We then take any over 5 (our upper bound) to be nulls.

```{r rate movies}
tblRatings<-expand.grid(tblMovies$movie_id,tblRaters$rater_id)
names(tblRatings) <- c("movie_id", "rater_id")
tblRatings$rating <-  sample(6, size = nrow(tblRatings), replace = TRUE)
tblRatings[which(tblRatings["rating"] > 5),]$rating <- NA
tail(tblRatings)
```

We then store and fetch our ratings

```{r}
dbBegin(conn)
dbExecute(conn,sqlAppendTable(conn, "movieratings", tblRatings,row.names = FALSE ))
dbCommit(conn)
tblRatings = dbReadTable(conn,"movieratings")
kable(head(tblRatings))
```

### Joining and getting summary data

If we want we can get some summary data for our ratings but first we must join them.

```{r}

tblAvg <- dbGetQuery(conn,"SELECT movie_name, avg(rating) as avgrating
   from movieratings INNER JOIN movies ON movieratings.movie_id = movies.movie_id  INNER JOIN raters ON movieratings.rater_id = raters.rater_id  GROUP BY movie_name ORDER BY movie_name")
kable(head(tblAvg))

```